### 2025-03-23 paper/main.tex
Notes, comments, tiny bibliographic formatting adjustments.

% typesetting issues
\newcommand{\documentname}{\textsl{Note}}
\newcommand{\sectionname}{Section}
\newcommand{\secref}[1]{\sectionname~\ref{#1}}
\setlength{\textwidth}{4.75in}
\setlength{\oddsidemargin}{3.25in}\addtolength{\oddsidemargin}{-0.5\textwidth}
\addtolength{\topmargin}{-0.20in}
\addtolength{\textheight}{1.40in}
\sloppy\sloppypar\raggedbottom\frenchspacing
\setstretch{1.08}
\pagestyle{myheadings}
\markboth{foo}{\scshape hainje {\footnotesize\&} hogg / useless formula for the area of a triangle}
% \markboth{foo}{\scshape hainje et al. / useless formula for the area of a triangle}

\titleformat*{\subsection}{\normalfont\normalsize\slshape}
\titlespacing*{\subsection}{0pt}{0pt}{0pt}

\usepackage{natbib}
\bibliographystyle{aasjournal}
\setcitestyle{aysep={}}
\setlength{\bibsep}{0pt plus 0.3ex}

\newcommand{\abs}[1]{|\,{#1}\,|}
\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}

\usepackage{xcolor}
\newcommand{\CH}[1]{{\color{blue} (CH says: #1)}}
\newcommand{\Hogg}[1]{{\color{violet} (Hogg says: #1)}}

@@

\renewcommand{\d}{\mathrm{d}}

\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}

\begin{document}\thispagestyle{empty}
\setcounter{secnumdepth}{0}

@@

\bigskip
\noindent
\textbf{Connor Hainje}\aOne\aSix{}
\&
\textbf{David W. Hogg}\aOne\aTwo\aThree%, and
%\textbf{Soledad Villar}\aFour\aFive

@@

Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
Using this fact, we have written code to perform a symbolic regression which recovers our result.\footnote{
    The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
}

In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
All the possible monomials are made of permutation-invariant factors of the form 
\begin{equation}
F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~.
\end{equation}
Such a factor has degree $m + n$.
The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
However, these monomials are not all linearly-independent;
the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
This leaves 28 linearly-independent monomials.

@@

\paragraph{Derivation of Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
The energy density $\mathcal{E} = d U / d^3 x$ is given by
\begin{equation}
    \mathcal{E} = \frac{\rho \, \Phi,}{2}~,
\end{equation}
where $\rho$ is the mass density and $\Phi$ is the gravitational potential \citep{LandauLifshitzFields}.
These are related by the Poisson equation
\begin{equation}
    \nabla^2 \Phi = 4 \pi G \, \rho~,
\end{equation}
which implies the following relationship between their Fourier transforms
\begin{equation}
    -k^2 \, \mathcal{F}[\Phi] = 4 \pi G \, \mathcal{F}[\rho]~,
\end{equation}
with $k^2 \equiv |\vec{k}|^2$.
Using the convolution theorem, we then have
\begin{align}
    \mathcal{F}[\mathcal{E}] 
    &= \frac{1}{2} \mathcal{F}[\rho \, \Phi] \nonumber\\
    &= \frac{1}{2} \, \frac{\mathcal{F}[\rho] \ast \mathcal{F}[\Phi]}{(2\pi)^3} \nonumber\\
    &= \frac{1}{2 \, (2\pi)^3} \, \mathcal{F}[\rho] \ast \left(
        -4 \pi G \, k^{-2} \, \mathcal{F}[\rho]
    \right)~.
    \label{eq:ConvolutionThm}
\end{align}
This form of the convolution theorem is proven below.
The Fourier transform of the mass density is
\begin{equation}
    \mathcal{F}[\rho](\vec{k})
    = \int \d^3 r
        \, e^{-i \, \vec{k} \cdot \vec{r}}
        \, \sum_{j=1}^{N}
        \, m_j
        \, \delta(\vec{r} - \vec{r}_j)
    = \sum_{j=1}^{N} \, m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}~,
\end{equation}
where we sum over $j$ and use $i=\sqrt{-1}$.
Note that the value of the Fourier transform at a wavenumber $\vec{k}$ is of the required form $\sum_i \phi(m_i, \vec{r}_i)$ for Deep Sets features.
Hence, the total energy is given in Deep Sets form by
\begin{align}
    \phi_{\vec{k}}(m_j, \vec{r}_j)
    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
    ~,
    \quad
    \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
    = \mathcal{F}[\rho](\vec{k})~,
    \nonumber\\
    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
    &= -\frac{G}{4 \pi^2} \,
        \int \d^3 r \,
        \mathcal{F}^{-1} \left[
            \mathcal{F}[\rho] \ast \left(
            k^{-2} \, \mathcal{F}[\rho] \right)
        \right]~.
\end{align}

>>>>

% page typesetting issues
\newcommand{\documentname}{\textsl{Note}}
\setlength{\textwidth}{4.75in}
\setlength{\oddsidemargin}{3.25in}\addtolength{\oddsidemargin}{-0.5\textwidth}
\addtolength{\topmargin}{-0.20in}
\addtolength{\textheight}{1.40in}
\sloppy\sloppypar\raggedbottom\frenchspacing
\setstretch{1.08}

% document formatting and sectioning issues
\pagestyle{myheadings}
\renewcommand{\paragraph}[1]{\par\addvspace{1.5ex}\noindent\textsl{#1}~---}
\newcommand{\secbreak}{\bigskip{\centering\footnotesize $\triangle~~~\triangle~~~\triangle$\par}\bigskip\noindent}
\titleformat*{\subsection}{\normalfont\normalsize\slshape}
\titlespacing*{\subsection}{0pt}{0pt}{0pt}
\markboth{foo}{\scshape hainje {\footnotesize\&} hogg / useless formula for the area of a triangle}
% \markboth{foo}{\scshape hainje et al. / useless formula for the area of a triangle}

% bibliography issues
\usepackage{natbib}
\bibliographystyle{aasjournal}
\setcitestyle{aysep={}}
\setlength{\bibsep}{0pt plus 0.3ex}

% math issues
\renewcommand{\d}{\mathrm{d}}
\newcommand{\abs}[1]{|\,{#1}\,|}

% issues
\usepackage{xcolor}
\newcommand{\CH}[1]{{\color{blue} (CH says: #1)}}
\newcommand{\Hogg}[1]{{\color{violet} (Hogg says: #1)}}

@@

\begin{document}\thispagestyle{empty}\setcounter{secnumdepth}{0}

@@

\medskip
\noindent
\textbf{Connor Hainje}\aOne\aSix{}
\textsl{and}
\textbf{David W. Hogg}\aOne\aTwo\aThree%, and
%\textbf{Soledad Villar}\aFour\aFive

@@

Note that a triangle defined by three points has one-half the area of the parallelogram spanned by the vectors $\vec{a} = \vec{r}_2 - \vec{r}_1$ and $\vec{b} = \vec{r}_3 - \vec{r}_1$.
Multiplying both sides by 4 reveals that \eqref{eq:result} is an integer polynomial of permutation-invariant factors for the squared area of the parallelogram.
Using this fact, we have written code to perform a symbolic regression that recovers our result.\footnote{
    The simple code is available for re-use under an open-source license at \url{https://github.com/davidwhogg/TriangleAreas}.
}

In this symbolic regression, we consider a linear combination of monomials, each of which can be written in Deep Sets form, but each of which has units of area-squared (position to the fourth power).
All the possible monomials are made of permutation-invariant factors of the form 
\begin{equation}
F_{mn} = \sum_{i=1}^{3} x_i^m \, y_i^n~;
\end{equation}
each such a factor has degree $m + n$.
The invariant monomials of degree $d$ are then products $\prod_j F_{m_j n_j}$ such that $\sum_j m_j + n_j = d$.
We generate all factors $F_{mn}$ with degree $1 \leq m + n \leq 4$, then form all products of these factors with total degree four.
However, these monomials are not all linearly-independent;
the five monomials made of a single degree-four factor ($F_{40}$, $F_{31}$, $F_{22}$, $F_{13}$, $F_{04}$) can be written as a linear combination of the other monomials and are thus removed.
This leaves 28 linearly-independent monomials.

@@

\paragraph{Derivation of Deep Sets energy} Consider a system of $N$ point particles with masses $m_i$ and positions $\vec{r}_i$.
The energy density $\mathcal{E} = d U / d^3 x$ is given by
\begin{equation}
    \mathcal{E} = \frac{\rho \, \Phi,}{2}~,
\end{equation}
where $\rho$ is the mass density and $\Phi$ is the gravitational potential \citep{LandauLifshitzFields}.
These are related by the Poisson equation
\begin{equation}
    \nabla^2 \Phi = 4 \pi G \, \rho~,
\end{equation}
which implies the following relationship between their Fourier transforms
\begin{equation}
    -k^2 \, \mathcal{F}[\Phi] = 4 \pi G \, \mathcal{F}[\rho]~,
\end{equation}
with $k^2 \equiv \abs{\vec{k}}^2$.
Using the convolution theorem, we then have
\begin{align}
    \mathcal{F}[\mathcal{E}] 
    &= \frac{1}{2} \mathcal{F}[\rho \, \Phi] \nonumber\\
    &= \frac{1}{2} \, \frac{\mathcal{F}[\rho] \ast \mathcal{F}[\Phi]}{(2\pi)^3} \nonumber\\
    &= \frac{1}{2 \, (2\pi)^3} \, \mathcal{F}[\rho] \ast \left(
        -4 \pi G \, k^{-2} \, \mathcal{F}[\rho]
    \right)~.
    \label{eq:ConvolutionThm}
\end{align}
This form of the convolution theorem is proven below.
The Fourier transform of the mass density is
\begin{equation}
    \mathcal{F}[\rho](\vec{k})
    = \int \d^3 r
        \, e^{-i \, \vec{k} \cdot \vec{r}}
        \, \sum_{j=1}^{N}
        \, m_j
        \, \delta(\vec{r} - \vec{r}_j)
    = \sum_{j=1}^{N} \, m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}~,
\end{equation}
where we sum over $j$ and use $i=\sqrt{-1}$.
Note that the value of the Fourier transform at a wavenumber $\vec{k}$ is of the required form $\sum_i \phi(m_i, \vec{r}_i)$ for Deep Sets features.
Hence, the total energy is given in Deep Sets form by
\begin{align}
    \phi_{\vec{k}}(m_j, \vec{r}_j)
    &=m_j \, e^{-i \, \vec{k} \cdot \vec{r}_j}
    ~,
    \quad
    \sum_{i=1}^{N} \phi_{\vec{k}}(m_i, \vec{r}_i)
    = \mathcal{F}[\rho](\vec{k})~,
    \nonumber\\
    U \, \big( \sum_{i=1}^{N} \, \phi_{\vec{k}}(m_i, \vec{r}_i) \big)
    &= -\frac{G}{4 \pi^2} \,
        \int \d^3 r \,
        \mathcal{F}^{-1} \left[
            \mathcal{F}[\rho] \ast \left(
            k^{-2} \, \mathcal{F}[\rho] \right)
        \right]~.
\end{align}
